---
title: "Homework 1: MPG Exploratory data analysis"
subtitle: Solution
author: "Satish G. Iyengar"
number-sections: false
format: html
editor: visual
---

## Load packages

```{r, warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(ggplot2)
library(urltools)
```


## Q1: What is a barplot? Draw barplot for the 'drv' variable using R or Python.
A bar plot is a graph that represents categorical data with rectangular bars. Each category is 
represented by a separate bar and the height of the bar represents the count of data points in that 
category.

### Read data

```{r}
dat_mpg <- mpg

```


```{r}
ggplot(dat_mpg, aes(x=drv, fill = drv)) + geom_bar() + 
  theme(legend.position = "top")
```

## What is a histogram? Draw histograms for highway miles per gallon - one for each drv type using R/Python. The histograms should be on the same scale for visual comparison.

A histogram is a type of chart that shows the frequency distribution of data points across a continuous range of numerical values. The values are grouped into bin or buckets that are arranged in consecutive order along the horizontal x-axis at the bottom of the chart. Each bin is represented by a vertical bar that sits on the x-axis and extends upward to indicate the number of data points within that bin [Ref: <https://www.techtarget.com/searchsoftwarequality/definition/histogram>].


```{r}
ggplot(dat_mpg, aes(x=hwy, fill = drv)) + 
  geom_histogram(alpha = 0.4, bins = 25, color = "black") + 
  theme(legend.position = "top")

```

## Q3: What is a box plot? Draw box plots for highway miles per gallon variable - one for each drv type in R/Python. The box plots should be on the same scale for visual comparison. What can you conclude from this box plot?

```{r}
ggplot(dat_mpg, aes(x= drv, y = hwy, fill = drv)) + geom_boxplot() +
  theme(legend.position = "top")
```

Based on the given data, the highway miles per gallon for front wheel drive cars have better fuel efficiency than both the all wheel drive and rear wheel drive cars.

