---
title: "HW 5: Detect Change in Variance"
author: "Satish G. Iyengar"
format: html
editor: visual
---

## 

Assume that we observe $x[n]$ for $n = 1, 2, ..., k$, where $x[n]'s$ are independent and identically distributed (i.i.d). That is, the probability density function of each $x[n]$ is the same, and there is no correlation between the $x[n]'s$. Further, it is also known that,

```{=tex}
\begin{eqnarray}
  H0: x[n] & ~ & Normal (mean = 0, variance = 1) \\
  H1: x[n] & ~ & Normal (mean = 0, variance = 1.5)
\end{eqnarray}
```
Design a detector and plot ROC assuming that we can collect 10 i.i.d samples for decision-making.

## 

## Simulate Data

```{r, warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)

```

```{r}
repM <- 10000 # number of MonteCarlo trials

k <- 10 # time points per decision

# H0 Model
muH0 <- 0
varH0 <- 1
sdH0 <- sqrt(varH0)

# H1 Model
muH1 <- 0
varH1 <- 1.5
sdH1 <- sqrt(varH1)

test_statistic_h0 <- NA
test_statistic_h1 <- NA
for(i in 1:repM){
  test_statistic_h0[i] <- var(rnorm(k, mean = muH0, sd = sdH0))
  test_statistic_h1[i] <- var(rnorm(k, mean = muH1, sd = sdH1))
}
dat_test_statistic <- 
  data.frame('test_stat' = test_statistic_h0, 'label' = 'H0') %>%
  bind_rows(data.frame('test_stat' = test_statistic_h1, 'label' = 'H1'))

```

### Probability Density of Test Statistics Under $H_0$ and $H_1$

```{r}
dat_test_statistic %>% 
  ggplot2::ggplot(., aes(x = test_stat, fill = label)) + 
  geom_density(alpha = 0.3)
```

## Plot Receiever Operating Characteristic Curve (ROC)

```{r}
aThreshold <- seq(0, 6, by = 0.05)
pd_est <- NA
pf_est <- NA
for(ith in 1:length(aThreshold)){
  th <- aThreshold[ith]
  pf_est[ith] <- sum(test_statistic_h0 > th)/length(test_statistic_h0)
  pd_est[ith] <- sum(test_statistic_h1 > th)/length(test_statistic_h1)
}

dat_roc <- 
  data.frame("threshold" = aThreshold, "pf_est" = pf_est, "pd_est" = pd_est)

```


```{r}
p1<-ggplot(dat_roc, aes(x=pf_est, y=pd_est)) + geom_line() +
  geom_point() + geom_abline(slope = 1, intercept = 0, col = "grey",
                             linetype = 'dashed', linewidth = 1)

```
