---
title: "Continuous Distribution"
author: Satish Iyengar
format: html
editor: visual
---

## Load packages

```{r, message=FALSE}
library(tidyverse)
library(plotly)
```


## Normal distribution

Let $X$ be a Gaussian distributed random variable, $X \sim \mathcal{N}(\mu, \sigma^2)$.
The probability density function (PDF) of $X$, $f_X(x)$, can be given as 
\begin{eqnarray}
  f_X(x) = \frac{1}{\sqrt(2 \pi \sigma^2)} \exp{\left(-\frac{1}{2\sigma^2}(x - \mu^2) \right)},  \hspace{5pt} x\in (-\infty, \infty)
  
\end{eqnarray}

To plot PDF of $X$, 

```{r}
x <- seq(-10, 10, by = 0.01)
mu <- 0
var <- 1

f <- dnorm(x, mu, sqrt(var))
dat_normpdf <- data.frame('x' = x, 'f' = f)
ggplot(dat_normpdf, aes(x = x, y = f)) + geom_line(size = 3) + 
  theme(axis.text = element_text(size = 20))
```

### Location of the distribution ($\mu$)


```{r}
MU    <- c(-5, 0, 5)
sigma <- sqrt(1)

dat_normpdf <- data.frame('x' = NA,'f' = NA, 'mu' = NA, 'sigma' = NA)
for(i in 1:length(MU)){
  mu <- MU[i]
  f <- dnorm(x, mu, sigma)
  tmp <- data.frame('x' = x,'f'=f, 'mu'= as.factor(mu), 'sigma' = sigma)
  dat_normpdf <- dat_normpdf %>%
    bind_rows(tmp)
  rm(tmp)
}

dat_normpdf <- dat_normpdf %>% drop_na()
```

```{r}
p <- ggplot(dat_normpdf, aes(x = x, y = f, color = mu)) + geom_line(size = 3)
p + theme(axis.text = element_text(size = 20))
```

### Spread of the distribution ($\sigma$)

```{r}
MU    <- c(0)
SIGMA <- round(c(sqrt(1), sqrt(5), sqrt(10)), digits = 2)

dat_normpdf <- data.frame('x' = NA,'f' = NA, 'mu' = NA, 'sigma' = NA)

for(i in 1:length(SIGMA)){
  mu <- MU
  sigma <- SIGMA[i]
  f <- dnorm(x, mu, sigma)
  tmp <- data.frame('x' = x,'f'=f, 'mu'= as.factor(mu), 'sigma' = as.factor(sigma))
  dat_normpdf <- dat_normpdf %>%
    bind_rows(tmp)
  rm(tmp)
}

dat_normpdf <- dat_normpdf %>% drop_na()
```

```{r}
p <- ggplot(dat_normpdf, aes(x = x, y = f, color = sigma)) + geom_line(size = 3)
p + theme(axis.text = element_text(size = 20))
```

## Uniform Distribution


Let $X$ be a Gaussian distributed random variable, $X \sim \mathcal{U}(a, b)$.
The probability density function (PDF) of $X$, $f_X(x)$, can be given as 
\begin{eqnarray}
  f_X(x) = \frac{1}{\left(b - a \right)},  \hspace{5pt} x\in (a, b)
\end{eqnarray}

To plot PDF of $X$, 

```{r}
# Define parameters of the uniform distribution U(a, b)
a <- 1
b <- 4

# x is a vector with different values of x
x <-  seq(0, 5, by = 0.001)
# we use the function dunif to compute pdf at all values of x
f <- dunif(x, min = a, max = b)
# store values of x and f in a dataframe
dat_unifpdf <- data.frame('x' = x, 'f' = f)

# Use ggplot to plot the density curve

p1 <- ggplot(dat_unifpdf, aes(x = x, y = f)) + 
  geom_line(color = "black", linewidth = 2) + 
  theme(text = element_text(size = 20)) + 
  xlab('x') + ylab('f(x)')

# Plot CDF 
Fx <- punif(x, min = a, max = b)
dat_unifcdf <- data.frame('x'=x,'Fx'=Fx)

p2 <- ggplot(dat_unifcdf, aes(x = x, y = Fx)) + 
  geom_line(color = "black", linewidth = 2) + 
  theme(text = element_text(size = 20)) + 
  xlab('x') + ylab('F(x)') + 
  geom_vline(xintercept = 2.5, linetype = 'dashed') + 
  geom_hline(yintercept = 0.5, linetype = 'dashed')

library(patchwork)
p1/p2


## Get 100 samples and plot histogram
N <- 100
xn <- runif(n=N, min = a, max = b)
hist(xn)

N <- 10000
xn <- runif(n=N, min = a, max = b)
hist(xn)

  
```

## Exerices

Suppose we have a cybersecurity system that detects network anomalies, and the response times of normal network requests are normally distributed with a mean of 100 milliseconds and a standard deviation of 20 milliseconds. Can we compute the probability that a network request takes between 80 milliseconds and 120 milliseconds to be processed? Use R/Python or the standard normal table (do web search for the standard normal table)

$$ X \sim \mathcal{N}(\mu = 100, \sigma ^2 = 400)$$
We want to compute $$P(80\leq X\leq 120) = P(X\leq 120) - P(X\leq 80) = F_X(120)-F_X(80). $$

```{r}
pnorm(120, mean = 100, sd = sqrt(400)) - pnorm(80, mean = 100, sd = sqrt(400)) 
```



```{r}
mu1 = 0
sd1 = sqrt(1)

x = rnorm(n = 1000, mean = mu1, sd = sd1)

hist(x)


x <- seq(-5, 5, by = 0.05)
f <- dnorm(x, mu1, sd1)

my_pdf <- data.frame('x' = x, 'f' = f)

ggplot(data = my_pdf, mapping = aes(x = x, y = f)) + geom_line() +
  geom_vline(xintercept = -1.96, linetype ='dashed') + 
  geom_vline(xintercept = 1.96, linetype ='dashed') 

my_probability <- pnorm(120, mean = mu1, sd = sd1) - pnorm(80, mean = mu1, sd = sd1)
  

```


## Compute $Pr(-1.96 < Z < 1.96)$

```{r}
pnorm(1.96, mean = 0, sd = 1) - 
  pnorm(-1.96, mean = 0, sd = 1)
  

```

```{r}

x <- seq(-5, 5, by =0.05)
F <- pnorm(x, mean = 0, sd =1) 
 
plot(x, F, type = 'l')

```



```{r}
qnorm(seq(0.001, 0.99991, by = 0.0005), mean = 0, sd = 1)


```

