---
title: "Flights Data Analysis"
format: 
  html:
    theme: journal
author: Satish Iyengar
---

## Questions

Find all flights that -

  - Had an arrival delay of two or more hours
  - Flew to Houston (IAH or HOU)
  - Were operated by United, American, or Delta
  - Departed in summer (July, August, and September)
  - Arrived more than two hours late, but didn’t leave late
  - Were delayed by at least an hour, but made up over 30 minutes in flight
  - Departed between midnight and 6 am (inclusive)



## Load packages

```{r, message=FALSE}
library(nycflights13)
library(tidyverse)
```

## Read and view data

```{r}
dat_flights <- nycflights13::flights 
dim(dat_flights)
dat_flights %>% head() %>% knitr::kable()
```

## Data description

- year, month, day: Date of departure
- dep_time, arr_time: Actual departure and arrival times (HHMM or HMM), local tz
- sched_dep_time, sched_arr_time: Scheduled departure and arrival times (HHMM or HMM), local tz

- dep_delay, arr_delay: Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.

- carrier: Two letter carrier abbreviation. See airlines to get name.

- flight: Flight number.

- tailnum: Plane tail number. See planes for additional metadata.

- origin, dest: Origin and destination. See airports for additional metadata.

- air_time: Amount of time spent in the air, in minutes.

- distance: Distance between airports, in miles.

- hour, minute: Time of scheduled departure broken into hour and minutes.

- time_hour: Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.

## Question 1: Flights with arrival delay of two or more hours

```{r}
dat_arr_delay_2hrs_or_more <- 
  dat_flights %>% dplyr::filter(arr_delay >= 120)
str(dat_arr_delay_2hrs_or_more) 
```

## Question 2: Flew to Houston (IAH or HOU)

```{r}
dat_to_IAH_HOU <- dat_flights %>% 
  dplyr::filter(dest %in% c('IAH','HOU'))
dat_to_IAH_HOU %>% select(dest) %>% table()
```

## Question 3: Were operated by United, American, or Delta

```{r}
nycflights13::airlines
```


```{r}
dat_AA_Delta_United <- 
  dat_flights %>% dplyr::filter(carrier %in% c('AA','DL','UA'))
dat_AA_Delta_United %>% dplyr::select(carrier) %>% table()
```


