---
title: 'Assignment 1: Exploratory data analysis'
output: html_document
date: "2023-09-18"
---


## Install packages

```{r, echo=TRUE}
# run this command if before loading. need to do this only once
# install.packages(tidyverse) 

```
# Set your working directory to where the CSV files are located
setwd("D:\School\BFOR 516 (Data Analytics for Cyber Sec) O")

## Load packages

```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(DT)

```



## Load the datasets
To read data from an external csv file use

```{r}
dat_mpg <- read.csv("D:/School/BFOR 516 (Data Analytics for Cyber Sec) O/mpg.csv")
```

## View data

```{r}
library(knitr)
DT::datatable(dat_mpg)
```

## 1) What is a barplot? Draw barplot for the 'drv' variable using R or Python.

```{r}
ggplot() + 
  geom_bar(data = dat_mpg, mapping = aes(x = drv, fill = drv)) + 
  theme(legend.position = 'none', text = element_text(size = 12), axis.text.x = element_text(angle = 90)) +
  labs(title = "Barplot for 'drv' Variable")
```


## 2) What is a histogram? Draw histograms for highway miles per gallon - one for each drv type using R/Python. The histograms should be on the same scale for visual comparison.
```{r}
ggplot() + 
  geom_histogram(data = dat_mpg, 
                 mapping = aes(x = hwy), 
                 fill="lightblue",
                 color="black", bins = 10) +
  facet_wrap(~drv, ncol=3) +
  theme(text = element_text(size = 20)) +
  labs (title = "Histograms of Highway MPG by Drive Type")
```



## 3)  What is a box plot? Draw box plots for highway miles per gallon variable - one for each drv type in R/Python. The box plots should be on the same scale for visual comparison. What can you conclude from this box plot?
```{r}
ggplot() + 
  geom_boxplot(data = dat_mpg, 
               mapping = aes(x = drv, y = hwy, fill= drv)) +
  theme(text = element_text(size = 12), legend.position = "none") +
  labs(title = "Box Plots of Highway MPG by Drive Type")
```


