---
title: "Lecture 2: Exploratory Data Analysis"
output: html_notebook
---


## Fuel Economy Dataset

## Install packages

```{r, echo=TRUE}
# run this command if before loading. need to do this only once
# install.packages(tidyverse) 

```

## Load packages

```{r, message = FALSE}
library(tidyverse)
library(knitr)
library(DT)

```

## Load data

```{r}
dat_mpg <- mpg

```

To read data from an external csv file use

```{r}
dat_mpg <- read_csv("mpg.csv")

```

## View data

```{r}
library(knitr)
DT::datatable(dat_mpg)
```

## Data description

```{r,}
# Type '?mpg' in RStudio console - what do you see?
```

- manufacturer: manufacturer name
- model: model name
- displ: engine displacement, in litres
- year: year of manufacture
- cyl: number of cylinders
- trans: type of transmission
- drv: the type of drive train
    - f: front-wheel drive
    - r: rear_wheel drive
    - 4: 4wd
- cty: city miles per gallon
- hwy: highway miles per gallon
- fl: fuel type
- class: "type" of car

## Data summary

```{r}
dat_mpg %>% summary()

```
## Data quality assessment 

- Missing values
- Outliers
- Domain-specific errors
- **Important**: Data governance and quality control 
  - Garbage in garbage out

### Count of missing values in 'hwy'

```{r}
dat_mpg %>% pull(hwy) %>% is.na() %>% sum()
```
### Count of missing values in all variables

```{r}
dat_mpg %>% summarise_all(., ~sum(is.na(.)))
```

## Visualization

### Manufacturer
```{r}
ggplot() + 
  geom_bar(data = dat_mpg, 
           mapping = aes(x = manufacturer, fill = manufacturer)) + 
  theme(legend.position = 'none',
    text = element_text(size = 12),
        axis.text.x = element_text(angle = 90),
    )
```

### Highway miles per gallon

```{r}
ggplot() + 
  geom_histogram(data = dat_mpg, 
                 mapping = aes(x=hwy), 
                 color="black", bins = 10) +
  theme(text = element_text(size = 20)) 
```

```{r}
dat_mpg %>% select(hwy) %>% summary()
```
### Highway mpg vs. Transmission type 

```{r}
dat_mpg %>% dplyr::group_by(trans) %>% summarise(n = n()) %>% kable()
```

```{r}
ggplot() + geom_bar(data = dat_mpg, 
                    mapping = aes(x=trans, fill = trans)) +
  theme(legend.position = "none", 
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 90))
```


#### Data transformation
```{r}
dat_mpg <- dat_mpg %>% 
  mutate(
    trans_binary = ifelse(startsWith(trans, "auto"), "auto", "manual")
  )

dat_mpg %>% dplyr::group_by(trans_binary) %>% 
  summarise(n = n())
```

```{r}
ggplot() + geom_bar(data = dat_mpg, 
                    mapping = aes(x=trans_binary, fill = trans_binary)) +
  theme(legend.position = "none", 
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 90))
```


```{r}
ggplot() + 
  geom_histogram(data = dat_mpg, 
                 mapping = aes(x = hwy, fill = trans_binary),
                 color = "black", 
                 alpha = 0.5) +
  theme(text = element_text(size = 12), 
        legend.position = "top") 
```
#### Boxplot 

```{r}
ggplot() + 
  geom_boxplot(data = dat_mpg, 
               mapping = aes(x = trans_binary, y = hwy, fill= trans_binary)) +
  theme(text = element_text(size = 12), legend.position = "top") 
```
```{r}
dat_mpg %>% group_by(trans_binary) %>% summarise(mean = mean(hwy), sd = sd(hwy), n = n())
```

## Exercise

- Highway mpg vs. drv
- Repeat analysis with city mpg
- Highway mpg vs. City mpg

### Submit by Sep 05, 2023

## Highway mpg vs. drv

```{r}
ggplot() + geom_bar(data = dat_mpg, 
                    mapping = aes(x=drv, fill = drv)) +
  theme(legend.position = "none", 
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 0))

```


```{r}
ggplot() + 
  geom_histogram(data = dat_mpg, 
                 mapping = aes(x = hwy, fill = drv),
                 color = "black", 
                 alpha = 0.5) +
  theme(text = element_text(size = 12), 
        legend.position = "top") 

```

```{r}
ggplot() + 
  geom_boxplot(data = dat_mpg, 
               mapping = aes(x = drv, y = hwy, fill= drv)) +
  theme(text = element_text(size = 12), legend.position = "top") 
```


