---
title: "Sampling Distribution"
format: html
editor: visual
---


```{r, message=FALSE}
library(tidyverse)
```

## Sampling from a Normally Distributed Population

$$ X \sim \mathcal{N}(\mu = 2, \sigma^2 = 4)$$
Sample from the above distribution 500 times for N = 2, N = 8 and N = 32, and examine sampling distributions.


```{r}
pop_mean <- 2
pop_sd <- 2

M = 500

N = 2
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rnorm(N, mean = pop_mean, sd = pop_sd))
}
xbar_df <- data.frame('xbar' = xbar, 'N' = as.character(N))

N = 8
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rnorm(N, mean = pop_mean, sd = pop_sd))
}
xbar_df <-xbar_df %>% 
  bind_rows(data.frame('xbar' = xbar, 'N' = as.character(N)))

N = 32
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rnorm(N, mean = pop_mean, sd = pop_sd))
}
xbar_df <-xbar_df %>% 
  bind_rows(data.frame('xbar' = xbar, 'N' = as.character(N)))

p <- ggplot(xbar_df, aes(x = xbar, fill = N)) + geom_density(alpha = 0.4)

p + theme(legend.position = 'top', 
          text = element_text(size=20)) + xlim(c(-4, 8)) + 
  geom_vline(xintercept = pop_mean, linetype = 'dashed')

```


- Sampling distributions for all N are centered around the population mean
- Variance of sample means decreases with increase in N

## Sampling from a non-Normally Distributed Population

$$ X \sim \mathcal{exp}(r=1)$$
Sample from the above distribution 500 times for N = 2, N = 8 and N = 32, and examine sampling distributions.


```{r}
r = 1
M = 500

N = 2
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rexp(N, rate = r))
}
xbar_df <- data.frame('xbar' = xbar, 'N' = as.character(N))

N = 8
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rexp(N, rate = r))
}
xbar_df <-xbar_df %>% 
  bind_rows(data.frame('xbar' = xbar, 'N' = as.character(N)))

N = 32
xbar = NA
for(i in 1:M){
  xbar[i] <- mean(rexp(N, rate = r))
}
xbar_df <-xbar_df %>% 
  bind_rows(data.frame('xbar' = xbar, 'N' = as.character(N)))

p <- ggplot(xbar_df, aes(x = xbar, fill = N)) + geom_density(alpha = 0.4)

p + theme(legend.position = 'top', 
          text = element_text(size=20)) + xlim(c(-2, 5)) +
  geom_vline(xintercept = 1/r, linetype = 'dashed')

```



## Central limit theorem

Let $X_1, X_2, \ldots , X_n$ denote a random sample of $n$ independent observations from a population with expected value $\mu$ and finite variance $\sigma ^{2}$, and let ${\displaystyle {\bar {X}}_{n}}$ denote the sample mean of that sample (which is itself a random variable). Then the limit as $n \to \infty$, the distribution of 
$${\displaystyle {\frac {{\bar {X}}_{n}-\mu }{\sigma _{{\bar {X}}_{n}}}},}$$ where 
${\displaystyle \sigma _{{\bar {X}}_{n}}={\frac {\sigma }{\sqrt {n}}},}$ is the standard normal distribution. That is,

$${\displaystyle {\frac {{\bar {X}}_{n}-\mu }{\sigma _{{\bar {X}}_{n}}}}} \sim \mathcal{N}(0, 1) $$

### Expected value and variance of sample mean

\begin{eqnarray}
 E[\bar{X}] & = & E\left[\frac{1}{N} \sum_{i=1}^N x_i\right] \\
 & = & \frac{1}{N}\left[\sum_{i=1}^N E[x_i]\right] \\
 & = & \frac{1}{N} N \mu \\
 & = & \mu
 \end{eqnarray}
 
 \begin{eqnarray}
 \sigma^2_{\bar{X}} = Var[\bar{X}] & = & Var\left[\frac{1}{N} \sum_{i=1}^N x_i\right] \\
 & = & \left(\frac{1}{N}\right)^2\left[\sum_{i=1}^N Var[x_i]\right] \\
 & = & \left(\frac{1}{N}\right)^2 N \sigma^2 \\
 & = & \frac{1}{N}\sigma^2
 \end{eqnarray}
 
 Therefore, the standard deviation, 
 $\sigma_{\bar{x}} = \frac{1}{\sqrt{N}} \cdot \sigma$
 

## Confidence Interval Estimation
 
 $${\displaystyle {\frac {{\bar {X}}_{n}-\mu }{\sigma _{{\bar {X}}_{n}}}}} \sim \mathcal{N}(0, 1) $$
 
 For a standard normal distribution we know that 
 $$ Pr(-1.96 \leq Z \leq 1.96) = 0.95$$.
 
 
That is,
 
 $$P\left({\displaystyle {-1.96 \leq \frac {{\bar {X}}_{n}-\mu }{\sigma _{{\bar {X}}_{n}}}}} \leq 1.96 \right) = 0.95$$
 
Now,

\begin{eqnarray}
-1.96 & \leq & \frac {{\bar{X}}_{n}-\mu }{\sigma_{{\bar{X}}_{n}}}\\
\implies -1.96 \cdot {\sigma_{{\bar{X}}_{n}}} & \leq & \bar{X}_n - \mu \\
\implies \mu & \leq & \bar{X}_n + 1.96 \cdot  {\sigma_{{\bar{X}}_{n}}}
\end{eqnarray}

and 

\begin{eqnarray}
\frac {{\bar{X}}_{n}-\mu }{\sigma_{{\bar{X}}_{n}}} & \leq & 1.96 \\
\implies \bar{X}_n - 1.96 \cdot  {\sigma_{{\bar{X}}_{n}}} \leq \mu
\end{eqnarray}

Thus, rewriting the probability equation we get,

\begin{eqnarray}
P\left(\bar{X}_n - 1.96 \cdot  {\sigma_{{\bar{X}}_{n}}} \leq \mu \leq \bar{X}_n + 1.96 \cdot  {\sigma_{{\bar{X}}_{n}}}\right) = 0.95
\end{eqnarray}


### Example
- Paper manufacturing facility
- Paper is expected to have a mean length of 11 inches and standard deviation of 0.02 inch.
- At periodic intervals, a sample is selected to determine if the mean paper length is 11 inches or something went wrong
- Say you selected 100 sheets at random and found that the mean length to be 10.998 inches. 
- What is the 95% confidence interval estimate for the population mean paper length?



\begin{eqnarray}
\bar{X} \text{ +/- } 1.96 \frac{\sigma}{\sqrt{N}} \\ 
= 10.998 \text{ +/- } (1.96) \cdot \frac{0.02}{\sqrt{100}} \\
= 10.998 +/- 0.0039
\end{eqnarray}

Thus, $10.9941 \leq \mu \leq 11.0019$

- Thus, with 95% confidence we can conclude that the population mean is between 10.9941 and 11.0019 inches.

#### How will you construct a 99% confidence interval?
